A fast algorithm for finding efficient decisions in multi-objective problems with black-box multiextremal functions

### 2022-01-10 2022/OM&S/paper.tex
Begining of Section 4

\section{Numerical experiments}

>>>>

\section{Computational experiments}

For efficiency assessment of the computational scheme described in the previous section the experiment with multiextremal multidimensional MCO benchmarks was conducted. These benchmarks were constructed on the base of multiextremal functions belonging to the class GKLS \cite{Gaviano2003} widely used in testing global optimization methods, i.e., the partial criteria $w_i(x), \; 1 \leq i \leq k,$ were generated as GKLS functions.

Along with the proposed technique these MCO benchmarks were solved by 5 metaheuristic methods. Two algorithms -- SMPSO \cite{Nebro2009} and OMOPSO \cite{Durillo2010} -- belong to the class of particle swarm methods and 3 others are genetic ones (NSGA-II \cite{Deb2002}, SPEA2 \cite{Zitzler2001}, IBEA \cite{Zitzler2004}). Implementation of these methods was taken from the library jMetalPy \cite{Benitez2019}. It should be noted that metaheuristic algorithms do not use convolutions to assess the Pareto set.

For estimating the effectiveness of solving the test MCO problem the hypervolume (HV) index  \cite{} was used as the quality criterion of the Pareto set approximations built by the methods. Moreover, besides comparing the efficiency of tested algorithms via their HVs, an ``ideal'' value of HV was computed by means of selection of non-dominated nodes of a dense uniform grid placed in the feasible domain.

In total, three groups of experiments have been performed. Each group contained a series of MCO problems of the same dimension and number of criteria. Since all GKLS functions have a global minimum value equal to $-1$, the partial  criteria $w_i(x), \; 1 \leq i \leq k,$  increased by a positive constant to ensure the requirements (\ref{positive}).

In all experiments the domain
\[
Q = \left\{x \in R^N : -1 \leq x_j \leq -1, 1 \leq i \leq N \right\}
\]
was considered as the feasible one.

Each problem was solved by all the methods including the grid technique and corresponding HVs were determined.  Metaheuristic methods because of their stochastic nature sometimes realize unsuccessful runs. In order to estimate better their efficiency each of these methods carried out 10 launches when solving a separate MCO problem and the average result was taken into account.

During all experiments the method ALR was used with parameters $r=5.6$, $\sigma = 15$, $T=3$. Accuracy $\varepsilon$ from the stopping condition (\ref{eq:23}) was $0.05$ in bi-criteria problems and $0.01$ in 4-criteria ones.

For correct comparison of the algorithms (except grid method, of course) their parameters influencing the stopping condition were chosen to provide approximately the same number of trials. After the completion of solving all the group problems, HVs of each method were averaged.

In the first experiment 100 two-dimensional bi-criteria problems have been taken. The contour plots of a convolution $\Phi_\lambda(x)$ from (\ref{conv_problem})  demonstrating its multiextremal and non-smooth relief are presented in Fig.~\ref{fig1}. For estimating the Pareto set by RALR-DA, in each MCO problem 50 convolutions (\ref{conv_problem}) corresponding to 50 values $\lambda$ uniformly distributed over interval $[0,1]$ have been minimized.

The Pareto front in one of test problems is shown in Fig.~\ref{fig2}.

The results of the experiment are presented in Tab.~\ref{tab1} and Fig.~\ref{fig3} where Pareto corresponds to the estimation obtained by the grid method.
